---
title: "pedoclimatic_pca"
author: "Léa Genty"
date: "2024-01-22"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(hrbrthemes)
library(viridis)
library(FactoMineR)
library(lubridate)
library(factoextra)
library(corrplot)
library(lme4)
library(MuMIn)
library(MASS)
library(AICcmodavg)
library(car)
library(RColorBrewer)
library(ggpubr)
library(ggsignif)
library(ggcorrplot)
```

# 1. Open data
```{r}
agroenvi <- read.csv(file="3. Pedoclim data/3.2. Clean pedoclimatic variables/Raw/Agroenvironnement.csv",sep=";",fileEncoding="latin1")
sol_var <- read.csv(file="3. Pedoclim data/3.2. Clean pedoclimatic variables/Raw/clean_soil_var.csv",sep=",") %>% 
  dplyr::select(-X)

id_code <- agroenvi %>% 
  dplyr::select(id_parcelle,code_parcelle,region_biogeo) %>% 
  distinct()

clim <- read.csv(file="3. Pedoclim data/3.2. Clean pedoclimatic variables/Raw/2023-07-12_Sol_climat.csv",sep=";") %>% 
  left_join(id_code) %>% 
  dplyr::select(id_parcelle,temp_moy_an,temp_range_an,  precip_an, precip_season) %>% 
  mutate(id_parcelle=as.character(id_parcelle))
```

# 2. Cleaning soil data
## 2.1. Keeping only old soil data for plots without measurements
```{r}
soil_old <- agroenvi %>% 
  dplyr::select(id_parcelle,Soil_Nitrogen,Soil_CEC,Soil_Clay_content,Soil_Ph_h2o,Soil_organic_carbon_content,Soil_Sand_content,Soil_Silt_content) %>% 
  distinct() %>% 
  filter(!id_parcelle %in% sol_var$id_parcelle) %>% 
  mutate(Soil_CEC=as.numeric(Soil_CEC),
         Soil_Clay_content=as.numeric(Soil_Clay_content),
         Soil_Nitrogen=as.numeric(Soil_Nitrogen),
         Soil_Ph_h2o=as.numeric(Soil_Ph_h2o),
         Soil_Sand_content=as.numeric(Soil_Sand_content),
         Soil_Silt_content=as.numeric(Soil_Silt_content),
         Soil_organic_carbon_content=as.numeric(Soil_organic_carbon_content))
```

## 2.2. Both datasets at the same unit and types of variables
```{r}
sol_new <- sol_var %>% 
  mutate(Soil_Sand_content=sables_fins+sables_grossiers,
         Soil_Silt_content = limons_grossiers+limons_fins) %>% 
  rename(
         Soil_Nitrogen=N_tot,
         Soil_CEC=CEC_a,
         Soil_Clay_content = argiles,
         Soil_organic_carbon_content = C_orga,
         Soil_Ph_h2o = pH
         ) %>% 
  dplyr::select(-sables_fins,-sables_grossiers,-limons_grossiers,-limons_fins,-CEC_b,-phosphore_mg_kg) %>% 
  mutate(Soil_Ph_h2o=Soil_Ph_h2o*10,
         Soil_Nitrogen= Soil_Nitrogen*100,
         Soil_organic_carbon_content=Soil_organic_carbon_content*10)  %>% 
  mutate(Soil_Clay_content=as.numeric(Soil_Clay_content),
         Soil_Silt_content=as.numeric(Soil_Silt_content),
         Soil_Sand_content=as.numeric(Soil_Sand_content))
```
Pour new :
C orga g/kg
1/1000
= 10 dg / kg
--> Faire x10

Azote Dumas g/kg 
1/1000
= 100 cg / kg
--> Faire x100

CEC mEq.kg

pour data old voir https://soilgrids.org/
C orga : dg/kg
1/10 000

Azote : cg/kg
1/100 000

CEC : mmol(c)/kg

1 g = 10 dg
1 g = 100 cg

## 2.3. Verifying old datas with correlations
```{r}
sol_new1 <- sol_var %>% 
  mutate(Soil_Sand_Content1=sables_fins+sables_grossiers,
         Soil_Silt_content1 = limons_grossiers+limons_fins) %>% 
  rename(
         Soil_Nitrogen1=N_tot,
         Soil_CEC1=CEC_a,
         Soil_Clay_content1 = argiles,
         Soil_organic_carbon_content1 = C_orga,
         Soil_Ph_h2o1 = pH
         ) %>% 
  dplyr::select(-sables_fins,-sables_grossiers,-limons_grossiers,-limons_fins,-CEC_b,-phosphore_mg_kg) %>% 
  mutate(Soil_Ph_h2o1=Soil_Ph_h2o1*10,
         Soil_Nitrogen1= Soil_Nitrogen1*100,
         Soil_organic_carbon_content1=Soil_organic_carbon_content1*10) %>% 
  mutate(id_parcelle=as.integer(id_parcelle))

sol_tot <- agroenvi %>% 
  filter(id_parcelle %in% sol_var$id_parcelle) %>% 
  left_join(sol_new1) %>% 
dplyr::select(id_parcelle,Soil_Sand_Content1,Soil_Silt_content1,Soil_Clay_content1,Soil_Nitrogen1,Soil_CEC1,Soil_organic_carbon_content1,Soil_Ph_h2o1,Soil_Sand_content,Soil_Silt_content,Soil_Clay_content,Soil_Nitrogen,Soil_CEC,Soil_organic_carbon_content,Soil_Ph_h2o)

cor.test(sol_tot$Soil_Nitrogen,sol_tot$Soil_Nitrogen1) #0.15 ***
cor.test(sol_tot$Soil_CEC,sol_tot$Soil_CEC1) # -0.12 ***
cor.test(sol_tot$Soil_Clay_content,sol_tot$Soil_Clay_content1) #0.49 ***
cor.test(sol_tot$Soil_Silt_content,sol_tot$Soil_Silt_content1) #0.63 ***
cor.test(sol_tot$Soil_Sand_content,sol_tot$Soil_Sand_Content1) #0.7 ***
cor.test(sol_tot$Soil_Ph_h2o,sol_tot$Soil_Ph_h2o1) # 0.6 ***
cor.test(sol_tot$Soil_organic_carbon_content,sol_tot$Soil_organic_carbon_content1) # 0.26 ***
```
Not good for Soil_Nitrogen, Soil_CEC and Soil_organic_carbon_content

Good for texture and pH

## 2.4. Adding new data and climatic data in the agroenvi dataset
```{r}
sol_tot <- soil_old %>% 
  mutate(id_parcelle=as.character(id_parcelle)) %>% 
  rbind(sol_new)

agroenvi_new <- agroenvi %>% 
  dplyr::select(-Soil_Nitrogen,-Soil_CEC,-Soil_Clay_content,-Soil_Ph_h2o,-Soil_organic_carbon_content,-Soil_Sand_content,-Soil_Silt_content) %>% 
  mutate(id_parcelle=as.character(id_parcelle)) %>% 
  left_join(sol_tot) %>% 
  left_join(clim) %>% 
  mutate(temp_moy_an = temp_moy_an/10,
         temp_range_an = temp_range_an/10)

#write.csv(agroenvi_new, file= "3. Pedoclim data/3.2. Clean pedoclimatic variables/Processed/Agroenvironnement_new.csv")
```
